### 2023-09-15 doc/_hierarchical_genotypes_results.qmd
several updates, including some of the comments by Griffin
continuing to make progress on paper

## Accounting for multiple barcodes per genotype via hierarchical models

Hierarchical models can also capture another experimental design in which
multiple barcodes map to the same or an equivalent genotype. As we will show,
this many-to-one mapping can improve the inference compared to the extreme cases
of inferring the fitness of each barcode independently or pooling the data of
all barcodes mapping to a single genotype. As shown in @fig-05(A), a small
modification to our model allows us to map the structure of our original model
to that of a hierarchical model with a fitness hyperparameter vector
$\underline{\theta}^G$, where $G$ is the number of genotypes in the datsets.

@@

![**Hierarchical model for multiple barcodes per genotype.** (A) Bayesian
network schematizing the statistical dependence between observations (filled
circles) and latent variables (empty circles). The barcodes belonging to
equivalent genotypes are connected via the hyper-parameters
$\underline{\theta}^G$. See `point to SI` for detailed definitions of each
variable. (B) Simulated dataset with 100 neutral lineages and 900 barcodes of
interest distributed among 90 genotypes. (C-E) Comparison between the inferred
and ground truth fitness values for a hierarchical model (C), a model where each
barcode is inferred independently (D), and a model where barcodes belonging to
the same genotype are pooled together (E). Gray error bars represent the 68\%
credible regions. (F) Empirical cumulative distribution function (ECDF) of the
absolute difference between the inferred mean and the ground truth fitness
values for all three models. (G) Examples of the posterior predictive checks for
all neutral lineages (upper left panels) and a subset of representative mutant
lineages. Shaded regions from light to dark represent the 95\%, 68\%, and 5\%
credible regions.](./figs/fig05){#fig-05}

>>>>

## Accounting for multiple barcodes per genotype via hierarchical models{#sec-genotypes}

Hierarchical models can also capture another experimental design in which
multiple barcodes map to the same or an equivalent genotype. As we will show,
this many-to-one mapping can improve the inference compared to the extreme cases
of inferring the fitness of each barcode independently or pooling the data of
all barcodes mapping to a single genotype. As schematized in @fig-05(A), a small
modification allows us to map the structure of our original model to that of a
hierarchical model with a fitness hyperparameter vector $\underline{\theta}^G$,
where $G$ is the number of genotypes in the dataset.

@@

![**Hierarchical model for multiple barcodes per genotype.** (A) Schematic
depiction of the hierarchical structure for multiple barcodes mapping to a
single genotype. A set of barcodes mapping to an equivalent genotype map to
"local" fitness values $s^{(b)}$ that are connected via a hyper-fitness
parameter for the genotype $\theta^{(g)}$. (B) Simulated dataset with 100
neutral lineages and 900 barcodes of interest distributed among 90 genotypes.
(C-E) Comparison between the inferred and ground truth fitness values for a
hierarchical model (C), a model where each barcode is inferred independently
(D), and a model where barcodes mapping to the same genotype are pooled
together (E). Gray error bars represent the 68\% credible regions. (F) Empirical
cumulative distribution function (ECDF) of the absolute difference between the
inferred mean and the ground truth fitness values for all three models. (G)
Examples of the posterior predictive checks for all neutral lineages (upper left
panels) and a subset of representative mutant lineages. Shaded regions from
light to dark represent the 95\%, 68\%, and 5\% credible
regions.](./figs/fig05){#fig-05}
